Theory of surface second-harmonic generation for semiconductors including effects of nonlocal operators

### 2014-11-21 paper-v-0.tex
done with abstract and intro

\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,11pt,preprint,letterpaper]{revtex4}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{ulem}
\usepackage{subfigure}
\usepackage{bm}
\usepackage[backref,pdffitwindow,colorlinks,linkcolor={blue},citecolor={red}]{hyperref}
%\usepackage{showlabels}
%\usepackage{showkeys}
\allowdisplaybreaks[1]%biutiful equation breaker!!!

@@

\begin{abstract}
We formulate a theoretical approach of surface second-harmonic generation 
from semiconductor surfaces based on the length gauge. 
Within the independent particle approximation 
the nonlinear second-order surface susceptibility tensor
$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$   
is calculated, including for the first time in one unique formulation
i) the scissors correction, needed to have the correct value of the
energy band gap,
ii) the derivation for the inclusion of the cut function, 
used to extract the surface nonlinear response,
and iii) the contribution of the nonlocal part of the 
pseudopotentials, routinely  used in \textit{ab initio} band structure calculations. 
First and third terms are described by spatially nonlocal quantum
mechanical operators and are fully taken into account in the present formulation. 
As a test-case of the approach we calculate 
$\chi^{xxx}(-2\omega;\omega,\omega)$ for the clean Si(001)$2\times 1$
reconstructed surface and find that our scheme to extract the surface
through the cut function
readily works and is exact. The effects of the scissors correction and
of the nonlocal part of the pseudopotentials are discussed for the
first time in surface nonlinear optics. 
The scissors correction shifts the spectrum to 
higher energies though the shifting is not rigid and mixes the 
$1\omega$ and $2\omega$ resonances, 
and has a strong influence in the line-shape. 
The 
effects of the nonlocal part of the pseudopotentials 
keeps the same line-shape of $|\chi^{xxx}_{2\times 1}(-2\omega;\omega,\omega)|$, but 
reduces its value
by 15-20\%.  
Therefore, the inclusion of the three aforementioned 
contributions is very important and 
makes our scheme unprecedented and opens the possibility to 
study SSHG with more versatility and providing more accurate results 
than previous approaches. 
\end{abstract}

@@

Finally, DFT-LDA calculations are often based on the use of pseudopotentials.
As it will be discussed 
in this article, the presence of a nonlocal part of the
pseudopotential introduces
corrections to
the momentum operator of the electron that have to be included with care in 
the SSH susceptibility. So far, this contribution has been neglected in the 
literature.

@@

The paper is organized as follows. In Sec. \ref{theory} we present the 
relevant steps for deriving the surface second-order susceptibility tensor
$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ within the 
length gauge formalism. This derivation includes the addition of the terms 
mentioned above that have been absent in previous works. Also, our 
$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ can be used 
for a layer-by-layer analysis if desired. In Sec. \ref{results} we show the 
results of this reformulation with a study of a clean Si(001) surface with a 
$2 \times 1$ reconstruction, 
proving the correctness of our approach with a special test-case,
never exploited before.
We compare results from before and after adding 
the different nonlocal contributions. Finally, in Sec. \ref{conc} we give 
our conclusions.

>>>>

%\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,11pt,preprint,letterpaper]{revtex4}
\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,reprint,letterpaper]{revtex4}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{ulem}
\usepackage{subfigure}
\usepackage{bm}
\usepackage[backref,pdffitwindow,colorlinks,linkcolor={blue},citecolor={red}]{hyperref}
%\usepackage{showlabels}
%\usepackage{showkeys}
\allowdisplaybreaks[1]%biutiful equation breaker!!!

@@

\begin{abstract}
We formulate a theoretical approach of surface second-harmonic generation from 
semiconductor surfaces based on the length gauge. We calculate the nonlinear 
second-order surface susceptibility tensor 
$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ within the 
independent particle approximation, and for the first time in one unique 
formulation we include i) the scissors correction, needed to have the correct 
value for the energy band gap, ii) the derivation for the inclusion of the cut 
function used to extract the surface nonlinear response, and iii) the 
contribution of the nonlocal part of the pseudopotentials that are routinely 
used in \textit{ab initio} band structure calculations. The first and third 
terms are described by spatially nonlocal quantum mechanical operators and 
are fully taken into account in the present formulation. As a test-case for 
this approach, we calculate $\chi^{xxx}(-2\omega;\omega,\omega)$ for a clean 
Si(001)$2\times 1$ reconstructed surface and find that this scheme for 
extracting the surface response through the cut function is exact and readily
implemented. The effects of the scissors correction and the nonlocal part of 
the pseudopotentials are discussed for the first time for surface nonlinear 
optics. The scissors correction non-rigidly shifts the spectrum to higher 
energies by mixing the $1\omega$ and $2\omega$ resonances and exerts a strong 
influence on the line-shape. The effects of the nonlocal part of the 
pseudopotentials keeps the same line-shape of 
$|\chi^{xxx}_{2\times 1}(-2\omega;\omega,\omega)|$ but reduces its value by 
15-20\%. Therefore, our scheme is unprecedented through the inclusion of 
these three contributions and also opens the possibility to study SSHG with 
more versatility and more accuracy than previous approaches. 
\end{abstract}

@@

Finally, DFT-LDA calculations are often based on the use of pseudopotentials.
As we will discuss later in this article, the presence of the nonlocal part 
of the pseudopotential introduces corrections to the momentum operator of 
the electron that have to be included with care into the SSH susceptibility. 
So far, this contribution has been neglected in the literature.

@@

The paper is organized as follows. In Sec. \ref{theory} we present the 
relevant steps for deriving the surface second-order susceptibility tensor
$\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ within the 
length gauge formalism. This derivation includes the addition of the terms 
mentioned above that have been absent in previous works. If desired, we can 
use $\chi^{\mathrm{a}\mathrm{b}\mathrm{c}}(-2\omega;\omega,\omega)$ for a 
layer-by-layer analysis. In Sec. \ref{results} we show the results of this 
reformulation with a study of a clean Si(001) surface with a $2 \times 1$ 
reconstruction that proves the validity of our approach with a special 
test-case that has never been implemented before. We compare results from 
before and after adding the different nonlocal contributions. Finally, in 
Sec. \ref{conc} we submit our conclusions.
